---
title: "Assignment 6"
author: "Your Name Here"
date: '2022-07-24'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# Add any packages you want in this chunk:
library(car)
library(gvlma)
```

We are going to look at the property prices for Orange County. From there we are going to try and predict the sales price by linear regression.

# Importing Data 

Bring in the data and make sure the data types are correct. If not, make the proper changes. The file is located within this project. *data/prop_prices_reduced.csv*

```{r}
housing <- read.csv("C:/Users/Ian Ah/Documents/GitHub/ECO5445/Assignment06/data/prop_prices_reduced.csv")
housing$pool <- factor(housing$pool) 
```


# Plotting

Plot histograms for all variables. Additionally, add scatterplots for the relationships between all quantitative variables.

```{r}
hist(housing$sale_def)
hist(housing$bed)
hist(housing$bath)
hist(housing$area_heated)
hist(housing$area)
hist(housing$dist_cbd)
hist(housing$dist_lakes)
hist(housing$pool)

plot(housing$sale_def, housing$area_heated)
plot(housing$sale_def, housing$bed)
plot(housing$sale_def, housing$bath)
plot(housing$sale_def, housing$area)
plot(housing$sale_def, housing$dist_cbd)
plot(housing$sale_def, housing$dist_lakes)
plot(housing$sale_def, housing$pool)

scatterplotMatrix(housing, spread = FALSE, lty.smooth = 2,
    main = "Scatterplot Matrix")
```

# Summary Statistics

Provide basic summary statistics for univariate analysis. Also, provide the correlation between all the quantitative variables.

```{r}
summary(housing)
cor(housing)
```

# Regression Analysis

Run a regression with all the variables included. Print results of the regression.

```{r}
Reg <- lm(sale_def ~ bed + bath + area_heated + area + dist_cbd + dist_lakes + pool, data = housing)
summary(Reg)
plot(Reg)
durbinWatsonTest(Reg)
gvmodel <- gvlma(Test_1)
summary(gvmodel)

```

Which of the variables tested significant at the 95% level? Looking at the results and answering outside of the chunk is sufficient.

Values that tested significant at the 95% level were sale_def, bed, bath, area_heated, area, dist_cbd, and pool. 

## Evaluating the model

As is, are any of the Gauss-Markov assumptions violated? If so, which ones? How can you fix the issues?

```{r}
#Assumption one is violated as many of the relationships between the variables are non-linear. For example, the relationship between sale_def and pools does not represent a linear relationship as the points plotted are 2 separate lines. 
```

## New Model

Based off of your findings in the previous section, make changes to the variables, the functional form, etc.

```{r}
housing$log_price <- (housing$sale_def/housing$area)
housing$log_lake <- (housing$dist_lakes^2)
housing$log_bath <- log(housing$bath)

Reg <- lm(log_price ~ bed + bath + area_heated + area + dist_cbd + dist_lakes + pool, data = housing)
summary(Reg)
Test_1 <- lm(sale_def ~ area_heated + bath + area + bed + dist_cbd + dist_lakes, data = housing)
summary(Test_1)
plot(Test_1)
plot(housing$sale_def, housing$log_lake)
```


# Prediction

Based on the following inputs, predict the deflated sales price:

- 2 bed
- 2 bath
- area_heated = 1223
- area = 9750
- dist_cbd = 19368
- dist_lakes = 490
- no pool

```{r}

```
